### 2016-06-29 publications/Journal/5-hotswap.tex
Rename files to fix the order again

Simply because a synchronization protocol is ideal at the start of the simulation, does not mean that it will still be ideal during the simulation.
It has been frequently shown, and repeated in the previous section, that model behaviour significantly influences the ideal synchronization protocol.
Contrary to many modelling formalisms, the DEVS formalisms makes it possible to model basically any kind of discrete event model.
As such, it is possible for the model to significantly change its behaviour throughout the simulation.

Defining the ideal synchronization protocol at the start of the simulation, when information about future model behaviour might be scarce, might therefore not offer the best possible performance.
In dxex, we not only make it possible to define the synchronization protocol to use, but also to change this decision throughout simulation.
To do this, all kernels are notified of the switch and they are forced to stop simulation using the current synchronization core.
When stopped, each kernel instantiates a new core that is provided with the simulation state of the previous core.
Simulation is then resumed with the new cores after the previous ones are destroyed.

As usual, switching imposes an overhead and should thus only be done if the benefits outweigh the induced overhead.
This overhead depends on the size of the model and the number of simulation cores.
For a simple model and a few cores, the overhead is less than a single second.

Although we currently only support manual switches between different synchronization protocols, this is not necessarily the case.
Ideally, a new component is added to the simulation kernel, which monitors model behaviour and simulation performance, and toggles between them automatically.
Our interface is thus augmented with the necessary bindings for such a component to be defined.
Also, our interface is augmented with an interface for statistics gathering and model behaviour analysis.
The implementation of such a component is currently left open: algorithms can be heavily based on machine learning or similar approaches.

@@

We add performance metrics in the simulation kernel, which logs relevant performance metrics and processes them for use in other components.
These metrics include the number of events created and destroyed, the number of inter and intra kernel events, the number of rollbacks, the measured lookahead, details of the Global Virtual Time (GVT) and Earliest Output Time (EOT) calculations, and basic information on the fairness between different simulation kernels.
With all these metrics, a component can get a fair view on both model and simulation kernel behaviour.

For example, if the actually seen lookahead is significantly higher than the defined lookahead, it might be interesting to switch to optimistic synchronization.
When the number of rollbacks is very high, conservative synchronization might be ideal.
And when neither of these two is an option, the simulation might just (temporarily) have to fall back to sequential simulation.

For performance reasons, statistics gathering is optional due to the imposed overhead.

\subsubsection{Visualization of Communication}
To provide some more insight in the models we used as benchmarks previously, we created a simple visualization of the simulation trace.
This trace visualizes the allocation of the model and all defined connections.
For each connection, the number of events transferred is annotated.
Examples are shown for the three benchmark models used before: Figure~\ref{fig:Queue_allocation}, Figure~\ref{fig:interconnect_allocation_parallel}, and Figure~\ref{fig:phold_allocation} show traces for the Queue, Interconnect, and PHOLD models respectively.

@@

Naturally, results similar to this are relevant information that can be used by the hotswapping component.

>>>>

Simply because a synchronization protocol is ideal at the start of the simulation, does not mean that it will still be ideal during the simulation.
It is well known, and repeated in the previous section, that model behaviour significantly influences the ideal synchronization protocol.
Contrary to many modelling formalisms, the DEVS formalisms makes it possible to model basically any kind of discrete event model.
As such, it is possible for the model to significantly change its behaviour throughout the simulation.

Defining the ideal synchronization protocol at the start of the simulation, when information about future model behaviour is scarce, might therefore not offer the best possible performance.
In dxex, we not only make it possible to define the synchronization protocol to use, but also to change this decision throughout simulation.
To do this, all kernels are notified of the switch and they are forced to stop simulation.
When stopped, each kernel instantiates a new core, with the new synchronization protocol, that is provided with the simulation state of the previous core.
Simulation is then resumed with the new cores after the previous ones are destroyed.

As usual, switching imposes an overhead and should thus only be done if the benefits outweigh the induced overhead.
This overhead depends on the size of the model and the number of simulation cores.
For a simple model and a few cores, the overhead is less than a second.

Although we currently only support manual switches between different synchronization protocols, this is not necessarily the case.
Ideally, a new component is added to the simulation kernel, which monitors model behaviour and simulation performance, and toggles between them automatically.
Our interface is augmented with the necessary bindings for such a decision component.
Also, our interface is augmented with an interface for statistics gathering and model behaviour analysis.
The implementation of such a component is currently left open, but we envision algorithms heavily based on machine learning.

@@

We add performance metrics in the simulation kernel, which logs relevant performance metrics and processes them for use in other components.
These metrics include the number of events created and destroyed, the number of inter- and intra-kernel events, the number of rollbacks, the measured lookahead, details of the Global Virtual Time (GVT) and Earliest Output Time (EOT) calculations, and information on the fairness between simulation kernels.
With all these metrics, a component can get a fair view on both model and simulation kernel behaviour.

For example, if the actually seen lookahead is significantly higher than the defined lookahead, it might be interesting to switch to optimistic synchronization.
When the number of rollbacks is excessively high, switching to conservative synchronization might be considered.
And when neither of these two is an option, the simulation might just (temporarily) have to fall back to sequential simulation.

\subsubsection{Visualization of Communication}
To provide more insight in our benchmark models, we created a simple visualization of their simulation trace.
This trace visualizes the allocation of the model and all defined connections.
For each connection, the number of events transferred is annotated.
Examples are shown for the three benchmark models used before: Figure~\ref{fig:Queue_allocation}, Figure~\ref{fig:interconnect_allocation_parallel}, and Figure~\ref{fig:phold_allocation} show traces for the Queue, Interconnect, and PHOLD models respectively.
Using this information, we notice that the Interconnect benchmark indeed has a lot of inter-kernel events.
Despite the similar structure, the PHOLD model does not have as many inter-kernel events.
These results are relevant information that can be used by the hotswapping component.

@@

### 2016-06-29 publications/Journal/8-conclusions.tex
Rename files to fix the order again

Future work is possible in several directions.
Firstly, our implementation of optimistic synchronization should be more tolerant to low-memory situations.
In its current state, simulation will simply halt with an out-of-memory error.
Having simulation control, which can throttle the speed of nodes that use up too much memory, has been shown to work in these situations~\cite{FujimotoBook}.
Faster GVT implementations~\cite{Fujimoto:1997:CGV:268403.268404,Bauer:2005:SND:1069810.1070159} might further help to minimize this problem.
Secondly, the runtime switching between synchronization protocols can be driven using machine learning techniques. The simulation engine is already capable of collecting data to inform such a process.
Thirdly, activity algorithms, as already implemented by PythonPDEVS, can also be implemented in dxex, to determine how they influence simulation performance.
Finally automatic allocation is possible by analysis of the connections between models. This information is already used in dxex to construct the dependency graph in conservative synchronization. A graph algorithm that distributes models while avoiding cycles in the resulting kernel topology could be used to offer a parallel speedup in either optimistic or conservative synchronization, or serve as a default parallel allocation scheme that can be improved by the user.

>>>>

Future work is possible in several directions.
First, our implementation of optimistic synchronization should be more tolerant to low-memory situations.
In its current state, simulation will simply halt with an out-of-memory error.
Having simulation control, which can throttle the speed of nodes that use up too much memory, has been shown to work in these situations~\cite{FujimotoBook}.
Faster GVT implementations~\cite{Fujimoto:1997:CGV:268403.268404,Bauer:2005:SND:1069810.1070159} might further help to minimize this problem.
Second, the runtime switching between synchronization protocols can be driven using machine learning techniques.
The simulation engine is already capable of collecting data to inform such a process, and is designed to listen for commands from an external component.
Third, automatic allocation might be possible by analysis of the connections between models.
This information is already used in dxex to construct the dependency graph in conservative synchronization.
A graph algorithm that distributes models, while avoiding cycles, could be used to offer a parallel speedup in either optimistic or conservative synchronization.
Similarly, it could serve as a default parallel allocation scheme that can be improved by the user.
